---
title: "Trabajo Grupal"
author: "Grupo 4"
date: "3/10/2021"
output: html_document
---

```{r Limpiar ambiente de trabajo, include=FALSE}
rm(list=ls())
```

#Librerias usadas

```{r Librerias, echo=TRUE, message=FALSE, warning=FALSE}
library(readr)
library(tidyverse)
library(reshape)
```

# Importacion de los datos

```{r Importacion de los datos, echo=TRUE, warning=FALSE}
incendios <- read_csv('http://www.dsi.uminho.pt/~pcortez/forestfires/forestfires.csv')
```
# Un vistazo a los datos

```{r Vistazo a los datos, echo=FALSE, warning=TRUE}
head(incendios) 
tail(incendios)
```

```{r Etiquetas para month y day, warning=FALSE, include=FALSE}
monthlevels <- c('jan', 'feb', 'mar', 'apr', 'may', 'jun', 'jul', 'aug',
                 'sep', 'oct', 'nov', 'dec')

monthlabels <- c('Enero', 'Febrero', 'Marzo', 'Abril', 'Mayo','Junio', 'Julio',
                 'Agosto', 'Septiembre', 'Octubre', 'Noviembre', 'Diciembre')

daylevels <- c('mon', 'tue', 'wed', 'thu', 'fri', 'sat', 'sun')

daylabels <- c('Lunes', 'Martes', 'Miercoles', 'Jueves', 'Viernes', 'Sabado', 'Domingo')
```

# Coercionar caracteres a factor

```{r Coercion de los datos a factor}
incendios <- incendios %>% mutate ( X = as.factor(X),
                                    Y = as.factor(Y),
                                    month = factor(month, levels = monthlevels,
                                                   labels = monthlabels),
                                    day = factor(day, levels = daylevels,
                                                 labels = daylabels)
                                   )

rm(list = c('daylabels', 'daylevels', 'monthlabels', 'monthlevels')) 
```

Chequeo que hayan quedado bien
```{r Vistazo a la estructura ya modificada, echo=TRUE}
glimpse(incendios)
```

# Contar NA en incendios

```{r Contar NA en Incendios, echo=TRUE}
apply(X = is.na(incendios), MARGIN = 2, FUN = sum) 
```
No hay datos faltantes

###### Analisis Exploratorio Univariado ##########




# Analisis Exploratorio Multivariado

##Frecuencia absoluta de incendios por mes por dia de la semana
 Creo un dataframe con las variables mes, dia y cantidad de incendios

```{r Dataframe MesXDia, echo=FALSE}

mesxdia <- incendios %>% {table(.$month, .$day)} %>% melt()

colnames(mesxdia) <- c('mes', 'dia', 'cant')

mesxdia
```
# Mapa de calor del dataframe mesxdia

```{r}
ggplot(data = mesxdia, aes(x=mes, 
                           y=dia, 
                           fill=cant)) + 
  geom_tile(color='black') +
  scale_fill_gradient(low = "white", high = "red") +
  guides(fill = guide_colourbar(title = "Cant. de Incendios"))
```

El grafico parece indicar que la cantidad de incendios varia mucho mas por el 
mes del anio que por el dia de la semana. Tambien los incendios por dia de la 
semana varian de igual forma a traves de los meses del anio

# Mapa de los incendios

Ya que X e Y son cordenadas espaciales, hare un mapa para visualizar donde 
se han concentrado los incendios

Hago la frecuencia absoluta de la caantidad de incendios por ubicacion
y los pongo en un data frame donde cada fila es una ubicacion

```{r Mapa}
mapa <- incendios %>% {table(.$X, .$Y)} %>% melt()
```

La columnas son la coordenada en X, la coordenada en Y y la cantidad de incendios
```{r Renombrar las columnas de Mapa}
colnames(mapa) <- c('coordX', 'coordY', 'cant')
```

Corrijo los tipos de datos del data frame
```{r coercion a factor}
mapa <- mapa %>% mutate(coordX = as.factor(coordX),
                coordY = as.factor(coordY),
                cant= as.integer(cant))
mapa
```

Grafico del 'mapa'
```{r Grafico de calor de Mapa}
ggplot(data=mapa, aes(x=coordX, y=coordY, fill=cant)) +  
  geom_tile(color='black') +
  scale_fill_gradient(low = "black", high = "red") +
  guides(fill = guide_colourbar(title = "Cant. de Incendios"))
```

# Informacion adjunta en el dataset

X - x-axis spatial coordinate within the Montesinho park map: 1 to 9
Y - y-axis spatial coordinate within the Montesinho park map: 2 to 9
month - month of the year: "jan" to "dec" 
day - day of the week: "mon" to "sun"
FFMC - FFMC index from the FWI system: 18.7 to 96.20
DMC - DMC index from the FWI system: 1.1 to 291.3 
DC - DC index from the FWI system: 7.9 to 860.6 
ISI - ISI index from the FWI system: 0.0 to 56.10
temp - temperature in Celsius degrees: 2.2 to 33.30
RH - relative humidity in %: 15.0 to 100
wind - wind speed in km/h: 0.40 to 9.40 
rain - outside rain in mm/m2 : 0.0 to 6.4 
area - the burned area of the forest (in ha): 0.00 to 1090.84

The Fine Fuel Moisture Code (FFMC) represents fuel moisture of forest litter fuels under the shade of a forest canopy. It is intended to represent moisture conditions for shaded litter fuels, the equivalent of 16-hour timelag. It ranges from 0-101. Subtracting the FFMC value from 100 can provide an estimate for the equivalent (approximately 10h) fuel moisture content, most accurate when FFMC values are roughly above 80.

The Duff Moisture Code (DMC) represents fuel moisture of decomposed organic material underneath the litter. System designers suggest that it is represents moisture conditions for the equivalent of 15-day (or 360 hr) timelag fuels. It is unitless and open ended. It may provide insight to live fuel moisture stress.

The Drought Code (DC), much like the Keetch-Byrum Drought Index, represents drying deep into the soil. It approximates moisture conditions for the equivalent of 53-day (1272 hour) timelag fuels. It is unitless, with a maximum value of 1000. Extreme drought conditions have produced DC values near 800.

The Initial Spread Index (ISI) is analogous to the NFDRS Spread Component (SC). It integrates fuel moisture for fine dead fuels and surface windspeed to estimate a spread potential. ISI is a key input for fire behavior predictions in the FBP system. It is unitless and open ended.
# Fin del adjunto